---
title: "Institutional or Electoral Preferences?\\newline Judges' Extended Calculus"
author: "JBrandon Duck-Mayr"
date: "`r gsub(' 0', ' ', x = format(Sys.Date(), '%B %d, %Y'))`"
fontsize: 12pt
titlepage: true
abstract: "Abstract (typically should be 150 words or less)"
graphics: yes
output:
    quack::article:
        includes:
            in_header: preamble.tex
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(bookdown.theorem.preamble = FALSE)
quack::attach_common_packages()
```

# Introduction

Some popular accounts of the US Supreme Court's decisions in *NFIB v. Sebelius* and *Dobbs v. Jackson Women's Health Organization* paint Chief Justice John Roberts as an ardent institutionalist, seeking a middle way to protect the Court's legitimacy; but what if there was another dynamic at play?
A strand of the judicial politics literature argues that politicians may offload the responsibility for pulling the trigger on controversial policies to the courts, either because they do not have the leeway to enact their preferred policies or to avoid electoral consequences [CITE CITE CITE].
Generally, in these analyses, the courts are treated as only considering the *direct* policy ramifications of their decisions.
I take a new approach and consider that judges might take into account the policies that politicians will be able to pass in the periods following a controversial decision by the courts... as we've seen with the aftermath of the *Dobbs* decision, such cases can have an impact on citizens' voting intentions [CITE].

# Literature Review

# The Model

The problem I am focusing on is a multimember court faced with an opportunity to adopt a controversial policy that a majority of its members want to adopt, and I restrict my attention to the bargaining problem among that majority.
That is, they must determine how far they are willing to go given the controversial nature of the policy; do they fully implement the controversial policy and risk electoral backlash against their ideological allies in the government, or do they avoid adopting it for fear of electoral backlash?
Or perhaps do they take a half-measure, balancing their desire to implement this policy and their concern for the policies in the legislature's domain in the next period?

So, I consider a set $J$ of the judges of the potential majority coalition to enact this policy; each judge $i$'s preferences can be represented by the following utility function:

\begin{equation}
U_i(p; \alpha_i, \beta_i, \varepsilon)
= p\alpha_i + \left( 1 - p^\varepsilon \right)\beta_i.
\end{equation}

Here $p \in [0, 1]$ represents the extent to which the court's opinion adopts the controversial policy and $\alpha_i > 0$ is the value of the policy to judge $i$, so that $p\alpha_i$ represents the policy gain from a choice of $p$ from $i$'s perspective.
However, $i$ also cares about the policies that will be enacted by the government in the next period.
Normalizing the value of the platform of $i$'s political opponents to $0$, $\beta_i$ represents the expected benefit of legislative policy in the next period---that is, the value of the platform of $i$'s copartisans weighted by the *a priori* probability that they are elected into government.
Then $p^\varepsilon$ represents the electoral penalty against $i$'s copartisans for a particular choice of $p$; $\varepsilon > 1$ governs how quickly the electoral cost approaches its maximum value as $p$ increases.
When $\varepsilon$ is large, the judges can partially implement the controversial policy without the public paying much mind, but when $\varepsilon$ is low, electoral penalties begin mounting with any move toward the controversial policy.
This relationship is illustrated in Figure \@ref(fig:epsilon).
<!-- Add footnote about former parameter \kappa -->

```{r epsilon, fig.cap = "Electoral cost of $p$ for different values of $\\varepsilon$", fig.width = 5, fig.height = 3}
parse_text = function(x) parse(text = x)
p = seq(from = 0, to = 1, by = 0.01)
low_e  = p^2
mid_e  = p^4
high_e = p^8
b = c(high_e, mid_e, low_e)
n = c("epsilon == 8", "epsilon == 4", "epsilon == 2")
g = rep(n, each = length(p))
d = data.frame(p, b, g)
v = setNames(c("solid", "dashed", "dotted"), n)
ggplot(data = d, mapping = aes(x = p, y = b, linetype = g)) +
    geom_line(size = 1) +
    xlab(bquote(p)) + ylab(bquote(p^epsilon)) +
    scale_linetype_manual(values = v, name = "", labels = parse_text) +
    scale_y_continuous(breaks = c(0, 1/2, 1)) +
    theme_bw() +
    theme(
        legend.key.width = unit(2, "line"),
        legend.text = element_text(size = 10)
    )
```

## Deterministic $\varepsilon$

Starting with the simplest case, if there is no uncertainty over $\varepsilon$, and a judge $i$ were able to decide the case alone, the resulting decision theoretic problem would be easy for the judge to solve.

::: {.proposition}

If there is no uncertainty over $\varepsilon$ and only one judge, the judge chooses $p=\min\left\{1, \left(\frac{\alpha_i}{\varepsilon\beta_i}\right)^{1 /(\varepsilon-1)}\right\}.$

:::

However, when dealing with a multimember court, the judges must find a solution they can all live with.
Since the judges' "outside option" is not enacting the controversial policy to any extent, it is worth $\beta_i$; however, they can only access this option unilaterally if their refusing to sign onto the majority coalition's opinion would deprive it of majority status.
This is the simplest case when we have multiple judges, so let's start there.

In this case, any judge $i$ will refuse to sign on to an opinion with $p \neq 0$ unless $p$ is such that

$$
p > \left(\frac{\beta_i}{\alpha_i}\right)^{1 - \varepsilon}.
$$

## Stochastic $\varepsilon$

## Stochastic $\varepsilon$ With Private Signals

# Data and Methods

# Results

# Conclusion

```{r child = "appendix.Rmd"}
```
