---
title: "Institutional or Electoral Preferences?\\newline Judges' Extended Calculus"
author: "JBrandon Duck-Mayr"
date: "`r gsub(' 0', ' ', x = format(Sys.Date(), '%B %d, %Y'))`"
fontsize: 12pt
titlepage: true
abstract: "Abstract (typically should be 150 words or less)"
graphics: yes
output: quack::article
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
quack::attach_common_packages()
```

# Introduction

Some popular accounts of the US Supreme Court's decisions in *NFIB v. Sebelius* and *Dobbs v. Jackson Women's Health Organization* paint Chief Justice John Roberts as an ardent institutionalist, seeking a middle way to protect the Court's legitimacy; but what if there was another dynamic at play?
A strand of the judicial politics literature argues that politicians may offload the responsibility for pulling the trigger on controversial policies to the courts, either because they do not have the leeway to enact their preferred policies or to avoid electoral consequences [CITE CITE CITE].
Generally, in these analyses, the courts are treated as only considering the *direct* policy ramifications of their decisions.
I take a new approach and consider that judges might take into account the policies that politicians will be able to pass in the periods following a controversial decision by the courts... as we've seen with the aftermath of the *Dobbs* decision, such cases can have an impact on citizens' voting intentions [CITE].

# Literature Review

# The Model

The problem I am focusing on is a multimember court faced with an opportunity to adopt a controversial policy that a majority of its members want to adopt, and I restrict my attention to the bargaining problem among that majority.
That is, they must determine how far they are willing to go given the controversial nature of the policy; do they fully implement the controversial policy and risk electoral backlash against their ideological allies in the government, or do they avoid adopting it for fear of electoral backlash?
Or perhaps do they take a half-measure, balancing their desire to implement this policy and their concern for the policies in the legislature's domain in the next period?

So, I consider a set $J$ of the judges of the potential majority coalition to enact this policy; each judge $i$'s preferences can be represented by the following utility function:

\begin{equation}
U_i(p; \alpha_i, \beta_i, \varepsilon, \kappa)
= p\alpha_i + \left( 1 - \kappa p^\varepsilon \right)\beta_i.
\end{equation}

Here $p \in [0, 1]$ represents the extent to which the court's opinion adopts the controversial policy and $\alpha_i > 0$ is the value of the policy to judge $i$, so that $p\alpha_i$ represents the policy gain from a choice of $p$ from $i$'s perspective.
However, $i$ also cares about the policies that will be enacted by the government in the next period.
Normalizing the value of the platform of $i$'s political opponents to $0$, $\beta_i$ represents the expected benefit of legislative policy in the next period---that is, the value of the platform of $i$'s copartisans weighted by the *a priori* probability that they are elected into government.
Then $1 - \kappa p^\varepsilon$ represents the electoral penalty against $i$'s copartisans for a particular choice of $p$; $\kappa \in (0, 1)$ represents the electoral cost of full implementation of the controversial policy (when $p = 1$, then $1 - \kappa p^\varepsilon = 1 - \kappa$) and $\varepsilon > 0$ governs how quickly the electoral cost approaches its maximum value as $p$ increases.
If $\varepsilon \in (0, 1)$, any movement toward the controversial policy is very quickly punished by the electorate, if $\varepsilon = 1$, the cost is linear in $p$, and if $\varepsilon > 1$, the electorate does not much mind *some* movement toward the controversial policy, but eventually the backlash begins to build quickly as $p$ increases; this relationship is illustrated in Figure \@ref(fig:epsilon).

```{r epsilon, fig.cap = "Electoral cost of $p$ for different values of $\\varepsilon$", fig.width = 5, fig.height = 3}
parse_text = function(x) parse(text = x)
p = seq(from = 0, to = 1, by = 0.01)
low_e  = p^(1/4)
mid_e  = p
high_e = p^4
b = c(low_e, mid_e, high_e)
n = c("epsilon == 1/4", "epsilon == 1", "epsilon == 4")
g = rep(n, each = length(p))
l = parse_text(c("0", "kappa / 2", "kappa"))
d = data.frame(p, b, g)
v = setNames(c("solid", "dashed", "dotted"), n)
ggplot(data = d, mapping = aes(x = p, y = b, linetype = g)) +
    geom_line(size = 1) +
    xlab(bquote(p)) + ylab(bquote(kappa * p^epsilon)) +
    scale_linetype_manual(values = v, name = "", labels = parse_text) +
    scale_y_continuous(breaks = c(0, 1/2, 1), labels = l) +
    theme_bw() +
    theme(legend.key.width = unit(2, "line"))
```

# Data and Methods

# Results

# Conclusion
